---
title: "Visualizing taxon from occurrence core"
output:
  html_document:
    df_print: paged
---

Recently I stumbled across this great R library [sunburstR](https://github.com/timelyportfolio/sunburstR) which could produce interactive sunburst. Thought this could be a good way to visualise taxa information for [biodiversity occurrence data](http://rs.tdwg.org/dwc/terms/)

The dataset used in this demo is SOMBASE PYCNOGONIDS cited as below.

## Dataset

British Antarctic Survey. SOMBASE PYCNOGONIDS. Occurrence Dataset https://doi.org/10.15468/qtm508 accessed on 2018-01-31.

## Load libraries

```{r message=FALSE, warning=FALSE}
library(robis)
library(sunburstR)
library(dplyr)
```

## Request occurrences using [robis](https://github.com/iobis/robis) 

```{r message=FALSE, warning=FALSE}
sombase <- occurrence(resourceid = 2296)
head(sombase)
```

## Create taxa sequence

`sunburstR` works with sequence. So let's first concatenate the taxa columns separated by "-".

```{r}
taxa <- data.frame(apply(sombase[,c("phylum", "class", "order", "family", "genus")], 1, paste, collapse="-"))
colnames(taxa) <- "taxa-seq"
head(taxa)
```

## Aggregate information and plot sunburst

We can group the `taxa-seq` using group_by and count the frequency of each distinct sequence, then pipe the output into sunburst.

```{r}
taxa %>% 
  group_by(`taxa-seq`) %>%
  summarise(uniq_count = n_distinct(`taxa-seq`)) %>%  # count the frequency of each distinct taxa-seq
  sunburst()
```

