---
title: "Bin occurrence records"
output: html_document
---

## Get data

```{r message=FALSE, warning=FALSE}
library(robis)
occ <- occurrence(resourceid = 2296)
dim(occ)
```

## Dataset

British Antarctic Survey. SOMBASE PYCNOGONIDS. Occurrence Dataset https://doi.org/10.15468/qtm508 accessed on 2018-02-11.

## Simple plot with [obistools](https://github.com/iobis/obistools)

```{r message=FALSE, warning=FALSE}
library(obistools)
plot_map(occ)
```

It's quick and nice to have a check how the distribution of the occurrence records look like. A lot of points are overlapping each other. It is difficult to tell how many points are there at certain locations. 

## Bin occurrences with [rbokeh](http://hafen.github.io/rbokeh/)

```{r message=FALSE, warning=FALSE}
library(rbokeh)
library(dplyr)
attach(occ)

figure(width = 900, height = 450) %>%
  ly_map("world", color = "#707070", alpha = 0.8) %>%  # grey map
  ly_hexbin(decimalLongitude, decimalLatitude, xbins = 100, shape = 0.2,
            alpha = 0.8, palette = "Spectral6", trans = log, inv = exp) %>% # colour by log count  
  theme_plot(background_fill_color = "black") %>%  # black background
  theme_grid(grid_line_alpha = 0)  # remove grid

```

## Polar projected map with binned occurrences

```{r}
library(ggplot2)
library(viridis)  # colorblind friendly 

world <- map_data("world")

ggplot() + 
  geom_bin2d(data = occ, aes(x = decimalLongitude, y = decimalLatitude), bins = 100) +  # bin data
  geom_path(data = world, aes(x = long, y = lat, group = group), colour = "#c0c0c0") +  # map
  scale_y_continuous(name = "latitude", breaks = (-2:2) * 30) + 
  scale_x_continuous(name = "longitude", breaks = (-4:4) * 45) + 
  coord_map("ortho", orientation = c(-90, 0, 0)) +  # orthographic projection from South Pole
  scale_fill_viridis(option = "viridis", trans = "log") + # log scale for bin count
  theme(panel.background = element_rect("black"),  # dark background
        panel.grid = element_blank(),  # remove panel grid
        axis.text.x = element_blank())  # remove x-axis value

```

In this example we plot map as line (no fill) on top of the bins, so that the bins do not block the view of map. Another option to do this is to adjust the `alpha` of the geom, but it does not look good in this case.
